---
title: "READ_ME"
author: "Aïssa Morin"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE) 
                    # cache = FALSE) 
```

```{r}

require(kableExtra)
require(tidyverse)
require(here)
require(ggplot2)
require(readr)

```

## Data preparation

N.B: Copy of the original data from Yolan are in Chapter_2_Dryness... -> documents --> Donnees_Yolan --> Donnee_Yolan_05-2023--> 'Data_Fat_analyses_Yolan'

Original dataset was modified and saved under 'Data_Fat_analyses_raw' in 'Chapter2.Dryness. ... '--> data --> 'Data_Fat_analyses_raw'

Modifications (manual because color coded) on xlxs file -->  column addition: 'collection_interval, values (i.e. Yolan color code)) :  

+ blue  --> early : sample collected within 6 days, 
+ yellow -->  intermediate : sample collected between 6 and 8 days (included ?), 
+ red --> late : sample collected after 8 days (should be exclude from analyses)

For the updated fat data (from 8th June 2023) I have also remove raws 288 --> 309 corresponding to sample that have not been through the whole drying process yet (i.e. not usable).

Saved as a csv file for R processing 

__Read data__

use function 'read_fat_data_raw()' to read csv file with raw fat data from Yolan (from may 2023) & 'read_fat_data_raw_2()' to read updated fat data (from 8th June 2023).


```{r}
# Load functions and deps ?
devtools::load_all()

#DATA PREPARATION ####

# Read data fat analyses raw dataset

#fat_data_raw <- read_fat_data_raw()

# Read updated data fat analyses raw dataset (from 8th June 2023)

fat_data_raw_2 <- read_fat_data_raw_2()

# read body condition of prey available in the landscape 
landsc_BC_data <- read_data_land_BC()


```


__Clean dataset__

Aim at cleaning, organising and only keep necessary info from fat raw data using the function 'clean_raw_fat_data'

The function :  

+ removes all 'mass_Jxx' columns, 
+ removes columns: "samples_marrow_fat_rate", "samples_marrow_fat_rate_without_residuals", "bone_marrow_fat_rate_mean", "bone_marrow_fat_rate_mean_without_residuals","estimated_bc","estimated_bc_without_residuals",  "comments" and any columns after,
+ removes all replicated samples tested after J-9, etc... /ex 'Hip_B_046-J12'
+ removes empty raws after 233rd row
+ removes raws corresponding to Buffalo_01 --> a test sample that wasn't harvested following the study design (but opportunistically) and  must, thus, be excluded because it wasn't
+ removes raws corresponding to hyena's data or with no information on predator species (NA)

The function include an option to save the cleaned dataset, the user must specified 'save = TRUE', and the filename under which the file must be saved. The cleaned dataset is saved under ./output/clean_fat_data/ 


```{r}

clean_fat_data <- clean_raw_fat_data(fat_data_raw_2,
                                     save = T,
                                     filename = 'cleaned_fat_data')
```


<!-- __Save cleaned dataset__ -->

<!-- Aim at saving cleaned data for fat analyses 'save_cleaned_data' -->

<!-- The function takes 2 arguments :    -->

<!-- +  _cleaned_data_ : cleaned raw data for fat analyses (after using clean_raw_fat_data function) -->
<!-- + _filename_ : the name we want the dataset to be saved under -->

<!-- The cleaned dataset is saved under ./output/clean_fat_data/  -->

<!-- <br> -->
<!-- <br> -->

